A Comparative study of Dask and Spark for neuroimaging pipelines

### 2019-06-28 latex/paper.tex
some comments on methods

\newcommand{\TG}[1]{\color{cyan}From Tristan: #1 \color{black}}
\newcommand{\MD}[1]{\color{magenta}From Mathieu: #1 \color{black}}

@@

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine. It
provides in-memory computing, data locality, and lazy evaluation. Spark's main data
structure, the Resilient Distributed Dataset (RDD)~\cite{RDD}, is a fault-tolerant
and a parallel collection of data elements. Dataframes, Spark's other popular API, is
an overlay of RDDs with similar performance features. Spark's core is written in
Scala, but it also has a Java, an R, and a Python API called PySpark. \MD{Sentence is
ambiguous? could think that all three APIs are named PySpark} We used PySpark, given
the popularity of Python in neuroscience, although the serialization cost of Python
objects to Scala might be a substantial source of overhead. We used Spark's default
scheduler, Spark standalone, although YARN~\cite{vavilapalli2013apache} and
Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler,
a process called the \emph{master} coordinates resources provisioned by
\emph{workers} on the cluster. The application is submitted to the \emph{driver} that
in turn requests workers to the master and dispatches jobs to them. A job is divided
into stages to be processed onto the workers. Each stage's operation is represented
as a high-level task in the computation graph. The Spark standalone scheduler uses a
FIFO (First-In-First-Out) job scheduling policy. Processing jobs in order helps in
increasing the page hit rate. As our applications only have one job, the LIFO
(Last-In-First-Out) policy used for the stages is more pertinent \MD{Find better
phrasing} \TG{refer to the paper (\MD{Not sure which one})}. The Spark standalone
scheduler has two execution modes: client mode, where the driver runs in a dedicated
process on the cluster, and cluster mode, where the driver runs in a worker. We used
client mode as PySpark is not available in cluster mode. We used Apache Spark v2.4.0.

\subsubsection{Dask} Dask is a Big Data engine that is becoming increasingly popular
in the scientific Python ecosystem. Dask has five main APIs and data structures:
Array, Bag, DataFrame, Delayed, and Futures. We used them all except DataFrame as it
is not adapted to the type of data we used. Like Spark, all Dask APIs provide
in-memory computing, data locality, lazy evaluation, and fault tolerance by mean of
lineage. Dask's core is purely in Python and bypass most of the
\href{https://docs.python.org/3/glossary.html#term-gil}{GIL} to provide parallelism.
We used Dask due to the common use of the scientific Python ecosystem in
neuroscience. Dask represents application tasks as Dask graphs, produced by a
user-facing API and executed by a scheduler. In Dask, operations generate multiple
tiny tasks in the computation graph allowing an easier representation of complex
algorithms. We used the \href{https://distributed.dask.org/en/latest/index.html}{Dask
Distributed} scheduler, although \href{https://github.com/dask/dask-yarn}{Dask-Yarn}
and \href{https://github.com/mrocklin/dask-mesos}{Dask-Mesos} is also available. In
the Dask Distributed scheduler, a process called \textit{dask-scheduler}
administrates the resources provided by \textit{workers} in the cluster. The
scheduler receives jobs from clients and assigns tasks to available workers. Like
Spark scheduler, Dask Distributed also use a LIFO policy, it is meant to reduce the
memory footprint of the application by finishing the computation of a branch in the
Dask graph before starting new ones. We used Dask v1.1.4.

In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects, similar to Spark's RDD. It offers a programming
abstraction similar to the \href{https://toolz.readthedocs.io/en/latest/}{PyToolz
library}. An \href{https://docs.dask.org/en/latest/array.html}{Array} is used for the
processing of large arrays. It provides a distributed clone of the popular NumPy
library. A \href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a
parallel composition of
\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas
Dataframes} used to process a large amount of tabular data.
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary tasks
that do not fit in the Array, DataFrame or Bag APIs. Finally,
\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to Delayed
as they support arbitrary tasks, but they operate in real-time rather than lazily.

>>>>

\newcommand{\TG}[1]{\color{cyan}From Tristan: #1 \color{black}}
\newcommand{\MD}[1]{\color{magenta}From Mathieu: #1 \color{black}}
\newcommand{\VHS}[1]{\color{green}From Valerie: #1 \color{black}}

@@

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine. It
provides in-memory computing, data locality, and lazy evaluation. Spark's main data
structure, the Resilient Distributed Dataset (RDD)~\cite{RDD}, is a fault-tolerant
and a parallel collection of data elements. Dataframes, Spark's API for tabular data, is
an overlay of RDDs with similar performance features \VHS{It's probably fine, but the SparkSQL Engine available for Scala and Java do bring an added peformance value}.\VHS{Spark has another API (Datasets). i know you used 'popular' for DataFrames, but it's not like Spark has a plethora of APIs. I think the APIs should be mentioned after the API languages such that you only list the APIs relevant to Python (RDD and Dataframes)} Spark's core is written in
Scala, but it also has a Java, an R, and a Python API called PySpark. \MD{Sentence is
ambiguous? could think that all three APIs are named PySpark} We used PySpark, given
the popularity of Python in neuroscience, although the serialization cost of Python
objects to Scala might be a substantial source of overhead. We used Spark's default
scheduler, Spark standalone, although YARN~\cite{vavilapalli2013apache} and
Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler,
a process called the \emph{master} coordinates resources provisioned by
\emph{workers} on the cluster. The application is submitted to the \emph{driver} that
in turn requests workers to the master and dispatches jobs to them\VHS{maybe say tasks instead of jobs, otherwise it sounds like 1 job/worker}. A job is divided
into stages to be processed onto the workers. Each stage's operation is represented
as a high-level task in the computation graph. The Spark standalone scheduler uses a
FIFO (First-In-First-Out) job scheduling policy. Processing jobs in order helps in
increasing the page hit rate. As our applications only have one job, the LIFO
(Last-In-First-Out) policy used for the stages is more pertinent \MD{Find better
phrasing} \TG{refer to the paper (\MD{Not sure which one})} \VHS{I wouldn't explicitly state LIFO here. It's a bit complicated to be frank. Basically both Spark and Dask opt to complete a chain of dependencies rather than start a bunch of ``new chains''. And in this context, LIFO is great for the page hit rate, so you might want to reword the prior sentence as well}. The Spark standalone
scheduler has two execution modes: client mode, where the driver runs in a dedicated
process on the cluster, and cluster mode, where the driver runs in a worker. We used
client mode as cluster mode in not available in PySpark. We used Apache Spark v2.4.0.

\subsubsection{Dask} Dask is a Big Data engine that is becoming increasingly popular
in the scientific Python ecosystem. Dask has five main APIs and data structures:
Array, Bag, DataFrame, Delayed, and Futures. All APIs were used in our experiments,
with the exception of Dask Dataframes, which is intended for use with tabular data.
Like Spark, all Dask APIs provide
in-memory computing, data locality, lazy evaluation, and fault tolerance by mean of
lineage. Dask's core is purely in Python and bypasses most of the
\href{https://docs.python.org/3/glossary.html#term-gil}{GIL} to provide parallelism.
We used Dask due to the common use of the scientific Python ecosystem in
neuroscience\VHS{If you add Dask Arrays to your experiments here, i'd mention that their
APIs might be particularly well suited for the processing of high resolution neuroimaging data}. Dask represents application tasks as Dask graphs, produced by a
user-facing API and executed by a scheduler. In Dask, operations generate multiple
tiny tasks in the computation graph allowing an easier representation of complex
algorithms. We used the \href{https://distributed.dask.org/en/latest/index.html}{Dask
Distributed} scheduler, although \href{https://github.com/dask/dask-yarn}{Dask-Yarn}
and \href{https://github.com/mrocklin/dask-mesos}{Dask-Mesos} is also available. In
the Dask Distributed scheduler, a process called \textit{dask-scheduler}
administrates the resources provided by \textit{workers} in the cluster. The
scheduler receives jobs from clients and assigns tasks to available workers. Like
Spark scheduler, Dask Distributed also use a LIFO policy, it is meant to reduce the
memory footprint of the application by finishing the computation of a branch in the
Dask graph before starting new ones. We used Dask v1.1.4.

In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects, similar to Spark's RDD. It offers a programming
abstraction similar to the \href{https://toolz.readthedocs.io/en/latest/}{PyToolz
library}. An \href{https://docs.dask.org/en/latest/array.html}{Array} is used for the
processing of large arrays. It provides a distributed clone of the popular NumPy
library. A \href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a
parallel composition of
\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas
Dataframes} used to process a large amount of tabular data.
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary tasks
that do not fit in the Array, DataFrame or Bag APIs. Finally,
\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to Delayed
as they support arbitrary tasks, but they operate in real-time rather than lazily.
\VHS{maybe mention which one use threads and which use processes}
